---
title: "CMOA - data preprocessing"
author: "Ung Lik Teng"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Importing Libraries
```{r}
library(readxl)
library(ggmap)
library(ggplot2)
library(dplyr)
library(stringr)
library(rJava)
library(xlsx)
library(rebus)
library(lettercase)
library(genderizeR)
library(Rserve)
library(mvoutlier)
library(tidyr)
```
##Importing the spreadsheet
```{r}
cmoa <- read_excel("cmoa.xlsx", sheet =1)
```
##Code the gender of the artist using genderizeR. I also use dplyr package to subset relevant columns to construct the artist table.
```{r}
cmoa_artist <- select(cmoa, `title`, `party_type`, `full_name`, `cited_name`, `role`, `nationality`) 
cmoa_artist
colnames(cmoa_artist)
table(cmoa_artist['party_type'])
table(cmoa_artist['role'])
table(cmoa_artist['full_name'])
cmoa_title <- distinct(cmoa_artist['title'])
table(cmoa_artist['title'])

givenNames = findGivenNames(as.character(full_artist$first_name), progress = FALSE)
what <- genderize(as.character(full_artist$first_name), genderDB = givenNames, progress = FALSE)
```
##Geocoding the birthplace and deathplace of the artist 
```{r}
#GETTING THE TRIMMED DOWN ARTIST DATASET 
cmoa.artist <- select(cmoa, `artist_id`, `party_type`, `full_name`, `cited_name`, `role`, `nationality`, `birth_date`, `death_date`, `birth_place`, `death_place` ) 
cmoa.artist[duplicated(cmoa.artist$artist_id),]
cmoa.artist[duplicated(cmoa.artist["full_name"]),]
cmoa.artist <- cmoa.artist[order(cmoa.artist$full_name),]
cmoa.artist_nd<-cmoa.artist[!duplicated(cmoa.artist$full_name),]
cmoa.artist_nd["n_photo"] <- (cmoa %>% group_by(full_name) %>% summarise(work_count =n()))[2]
head(cmoa.artist_nd)

#BIRTHPLACE AND DEATHPLACE 
cmoa.art.bp <- select(cmoa.artist_nd, `full_name`, `nationality`, `birth_place`)
head(cmoa.art.bp)
cmoa.art.bp <- cmoa.art.bp[!is.na(cmoa.art.bp$birth_place),]
cmoa.art.bp[c("bp_lon", "bp_lat")] <- geocode(as.matrix(cmoa.art.bp$birth_place))
cmoa.art.bp <- filter(cmoa.art.bp, `birth_place` != "|")

cmoa.art.dp <- select(cmoa.artist_nd, `full_name`, `nationality`, `death_place`)
cmoa.art.dp <- cmoa.art.dp[!is.na(cmoa.art.dp$death_place),]
cmoa.art.dp[c("dp_lon", "dp_lat")] <- geocode(as.matrix(cmoa.art.dp$death_place))
cmoa.art.dp <- filter(cmoa.art.dp, `death_place` != "|")

geocodeQueryCheck(userType = "free")
```
##Some photographs have more than 1 creator/artist. To analyze only the photograph, I use stringr and regular expression to split the photographer and add it to the row below. This is done for both birthplace subset and deathplace subset 
```{r}
#Split photographer
cmoa.art.bp[c("photographer_1", "photographer_2")] <- str_split(as.matrix(cmoa.art.bp["full_name"]), pattern = fixed("|"), simplify = T)
cmoa.art.bp["full_name"] <- NULL
cmoa.art.bp[c("nationality_1", "nationality_2")] <- str_split(as.matrix(cmoa.art.bp["nationality"]), pattern = fixed("|"), simplify = T)
cmoa.art.bp["nationality"] <- NULL
cmoa.art.bp[c("birth_place_1", "birth_place_2")] <- str_split(as.matrix(cmoa.art.bp["birth_place"]), pattern = fixed("|"), simplify = T)
cmoa.art.bp["birth_place"] <- NULL
cmoa.art.bp[c("bp_lon_1", "bp_lat_1")] <- geocode(as.matrix(cmoa.art.bp$birth_place_1))
cmoa.art.bp[c("bp_lon_2", "bp_lat_2")] <- geocode(as.matrix(cmoa.art.bp$birth_place_2))
cmoa.art.bp[c("dp_lon","dp_lat")] <- NULL
write.xlsx(cmoa.art.bp, "cmoa_artist_birthplace.xlsx", sheetName="Sheet1", row.names = FALSE)

names(cmoa.art.bp)
bp1 <- select(cmoa.art.bp, `photographer_1`, `nationality_1`, `birth_place_1`, `dp_lon_1`, `dp_lat_1`)
bp2 <- select(cmoa.art.bp, `photographer_2`, `nationality_2`, `birth_place_2`, `dp_lon_2`, `dp_lat_2`)
names(bp1) <- c("photographer" , "nationality", "birth_place", "bp_lon", "bp_lat")
bp_combined <- t(cbind(t(bp1[,]), t(bp2[2:188,])))
bp_combined <- as.data.frame(bp_combined)
bp_combined <- filter(bp_combined, !(bp_combined$photographer == ""))

cmoa.art.dp["full_name"] <- NULL
cmoa.art.dp["full_name_1"] <- NULL
cmoa.art.dp[c("nationality_1", "nationality_2")] <- str_split(as.matrix(cmoa.art.dp["nationality"]), pattern = fixed("|"), simplify = T)
cmoa.art.dp["nationality"] <- NULL
cmoa.art.dp[c("death_place_1", "death_place_2")] <- str_split(as.matrix(cmoa.art.dp["death_place"]), pattern = fixed("|"), simplify = T)
cmoa.art.dp["death_place"] <- NULL
cmoa.art.dp[c("dp_lon_1", "dp_lat_1")] <- geocode(as.matrix(cmoa.art.dp$death_place_1))
cmoa.art.dp[c("dp_lon_2", "dp_lat_2")] <- geocode(as.matrix(cmoa.art.dp$death_place_2))
cmoa.art.dp[c("dp_lon","dp_lat")] <- NULL
write.xlsx(cmoa.art.dp, "cmoa_artist_deathplace.xlsx", sheetName="Sheet1", row.names = FALSE)

names(cmoa.art.dp)
dp1 <- select(cmoa.art.dp, `photographer_1`, `nationality_1`, `death_place_1`, `dp_lon_1`, `dp_lat_1`)
dp2 <- select(cmoa.art.dp, `photographer_2`, `nationality_2`, `death_place_2`, `dp_lon_2`, `dp_lat_2`)
names(dp1) <- c("photographer" , "nationality", "death_place", "dp_lon", "dp_lat")
dp_combined <- t(cbind(t(dp1[,]), t(dp2[2:188,])))
dp_combined <- as.data.frame(dp_combined)
dp_combined <- filter(dp_combined, !(dp_combined$photographer == ""))

bp_combined <- filter(bp_combined, !(duplicated(bp_combined$photographer)))
dp_combined <- filter(dp_combined, !(duplicated(dp_combined$photographer)))
which(duplicated(bp_combined$photographer))
which(duplicated(dp_combined$photographer))
```
##Do the same for artist's birthdate and deathdate
```{r}
names(cmoa.artist_nd)
cmoa.art.bd.dd <- select(cmoa.artist_nd, `full_name`, `birth_date`, `death_date`, `nationality`, `party_type`, `n_photo`)
head(cmoa.art.bd.dd)
names(cmoa.art.bd.dd)

cmoa.art.bd.dd[c("photographer_1", "photographer_2")] <- str_split(as.matrix(cmoa.art.bd.dd["full_name"]), pattern = fixed("|"), simplify = T)
cmoa.art.bd.dd["full_name"] <- NULL
cmoa.art.bd.dd[c("nationality_1", "nationality_2")] <- str_split(as.matrix(cmoa.art.bd.dd["nationality"]), pattern = fixed("|"), simplify = T)
cmoa.art.bd.dd["nationality"] <- NULL
cmoa.art.bd.dd[c("birth_date_1", "birth_date_2")] <- str_split(as.matrix(cmoa.art.bd.dd["birth_date"]), pattern = fixed("|"), simplify = T)
cmoa.art.bd.dd["birth_date"] <- NULL
cmoa.art.bd.dd[c("death_date_1", "death_date_2")] <- str_split(as.matrix(cmoa.art.bd.dd["death_date"]), pattern = fixed("|"), simplify = T)
cmoa.art.bd.dd["death_date"] <- NULL
cmoa.art.bd.dd[c("party_type1", "party_type2")] <- str_split(as.matrix(cmoa.art.bd.dd["party_type"]), pattern = fixed("|"), simplify = T)
cmoa.art.bd.dd["party_type"] <- NULL
cmoa.art.bd.dd["n_photo_2"]<- numeric(438)
cmoa.art.bd.dd[!(cmoa.art.bd.dd$photographer_2 == ""),"n_photo_2"]<-  cmoa.art.bd.dd[!(cmoa.art.bd.dd$photographer_2 == ""), 'n_photo' ]



bd.dd1 <-select(cmoa.art.bd.dd, `photographer_1`, `n_photo`, `nationality_1`, `party_type1`, `birth_date_1`, `death_date_1`)
bd.dd2 <- select(cmoa.art.bd.dd, `photographer_2`, `n_photo_2`, `nationality_2`, `party_type2`, `birth_date_2`, `death_date_2`)
names(bd.dd1) <- c("photographer" , "n_photo", "nationality", "party_type", "birth_date", "death_date")
bd.dd_combined <- t(cbind(t(bd.dd1[,]), t(bd.dd2[2:438,])))
bd.dd_combined <- as.data.frame(bd.dd_combined)
bd.dd_combined <- filter(bd.dd_combined, !(bd.dd_combined$photographer == ""))

pattern.bddd <- exactly(one_or_more(DGT))
bd_renew <- as.Date(as.numeric(str_subset(bd.dd_combined$birth_date, pattern = pattern.bddd)), origin = "1899-12-30")
index_bd <- which(str_detect(bd.dd_combined$birth_date, pattern = pattern.bddd))
bd.dd_combined$birth_date <- as.character(bd.dd_combined$birth_date)
bd.dd_combined[index_bd, 'birth_date'] <- as.character(bd_renew)

dd_renew <- as.Date(as.numeric(str_subset(bd.dd_combined$death_date, pattern = pattern.bddd)), origin = "1899-12-30")
index_dd <- which(str_detect(bd.dd_combined$death_date, pattern = pattern.bddd))
bd.dd_combined$death_date <- as.character(bd.dd_combined$death_date)
bd.dd_combined[index_dd, 'death_date'] <- as.character(dd_renew)

sum(is.na(bd.dd_combined$birth_date))
sum(is.na(bd.dd_combined$death_date))

first_four <- dgt(4)
bd_renew2 <- str_extract(bd.dd_combined$birth_date, pattern = first_four)
bd.dd_combined['birth_year'] <- as.character(bd_renew2)

dd_renew2 <- str_extract(bd.dd_combined$death_date, pattern = first_four)
bd.dd_combined['death_year'] <- as.character(dd_renew2)
bd.dd_combined$n_photo <- as.numeric(bd.dd_combined$n_photo)
bd.dd_combined <- filter(bd.dd_combined, !(duplicated(bd.dd_combined$photographer)))

cmoa.art.bd.dd$birth_date <- as.Date(cmoa.art.bd.dd$birth_date, format = "%d/%m/%Y")
cmoa.art.bd.dd$birth_date
which(duplicated(bd.dd_combined$photographer))
```
##The Creation Date column of the data contains a lot of noise such unwanted "." or "/".
##For this dataset, we are only interested in the year in which the photograph is created, hence we will clean the creation date column to standardize it to four digit years
```{r}
cmoa_date <- select(cmoa, `title`, `creation_date`, `creation_date_earliest`, `creation_date_latest`)
sum(is.na(cmoa_date$creation_date))
sum(is.na(cmoa_date$creation_date_earliest))
sum(is.na(cmoa_date$creation_date_latest))
cmoa_date[is.na(cmoa_date$creation_date_earliest), c(2,3,4)]

#this code chunk below remove the empty year, which is represented by "c."
c_dot <- "c" %R% DOT %R% SPC
sub_c_dot <- str_subset(cmoa_date$creation_date, pattern = c_dot) 
index_cdot <- which(str_detect(cmoa_date$creation_date, pattern = c_dot))
cmoa_date[index_cdot, 'creation_date'] <- str_replace(sub_c_dot, pattern = c_dot, replacement = "")

#some rows also show the printed years, we want the creation date hence we remove the printed years 
printed <- SPC %R% "printed"
sub_printed <- str_subset(cmoa_date$creation_date, pattern = printed)
index_printed <- which(str_detect(cmoa_date$creation_date, pattern = printed))
str_extract(sub_printed, pattern = (dgt(4)))
cmoa_date[index_printed, 'creation_date'] <- str_extract(sub_printed, pattern = (dgt(4)))

#We replace the NA values with the latest creation date  
cmoa_date[is.na(cmoa_date$creation_date), 'creation_date'] <- cmoa_date[is.na(cmoa_date$creation_date), "updated_latest_year"]

#Some creation date shows a range from year-year. I take the mean of the two years
xxxx <- dgt(4) %R% "-" %R% dgt(4)
xxxx_split <- str_split(xxxx_subset, pattern = "-")
xxxx_subset <- str_subset(cmoa_date$creation_date, pattern = xxxx)
numeric_xxxx <- sapply(xxxx_split, as.numeric, simplify = F)
mean_xxxx <- sapply(numeric_xxxx, mean, simplify = F)
round_xxxx <- sapply(mean_xxxx, round)
index_xxxx <- which(str_detect(cmoa_date$creation_date, pattern = xxxx))
cmoa_date[index_xxxx, 'creation_date'] <- round_xxxx

#there are years in the format of 1960s, I use randomization to estimate a year 
early_ies <- "early" %R% SPC %R% (dgt(4)) %R% "s"
late_ies <- "late"  %R% SPC %R% (dgt(4)) %R% "s"
early_set <- str_subset(cmoa_date$creation_date, patter = early_ies)
late_set <- str_subset(cmoa_date$creation_date, patter = late_ies)

set.seed(19292)
early_sample <- sample(0:4, size = length(early_set), replace = TRUE)
late_sample<- sample(5:9, size = length(late_set), replace = TRUE)

early_set <- str_replace(early_set, pattern = 0 %R% "s", replacement = as.character(early_sample))
late_set <- str_replace(late_set, pattern = 0 %R% "s", replacement = as.character(late_sample))
early_set <- str_replace(early_set, pattern = "early" %R% SPC, replacement = "")
late_set <- str_replace(late_set, pattern = "late" %R% SPC, replacement = "")
index_early <- str_detect(cmoa_date$creation_date, patter = early_ies)
index_late <- str_detect(cmoa_date$creation_date, patter = late_ies)

cmoa_date[index_early, 'creation_date'] <- early_set
cmoa_date[index_late, 'creation_date'] <- late_set

#1960s no dash
ies <- START %R% dgt(4) %R% "s" %R% END
ies_set <- str_subset(cmoa_date$creation_date, patter = ies)
set.seed(123408)
ies_sample <- sample(0:9, size = length(ies_set), replace = TRUE)
index_ies <- str_detect(cmoa_date$creation_date, patter = ies)

ies_set <- str_replace(ies_set, pattern = 0 %R% "s", replacement = as.character(ies_sample))
cmoa_date[index_ies, 'creation_date'] <- ies_set

#1960s - 
ies_dash <- 0 %R% "s"
iesdash_set <- str_subset(cmoa_date$creation_date, patter = ies_dash)

set.seed(1234123)
iesd_sam1 <- sample(0:9, length(iesdash_set), replace = T)

set.seed(59204)
iesd_sam2 <- sample(0:9, length(iesdash_set), r = T)

iesdash_set <- str_replace(iesdash_set, pattern = 0 %R% "s", replacement = as.character(iesd_sam1))
iesdash_set <- str_replace(iesdash_set, pattern = 0 %R% "s" %R% END, replacement = as.character(iesd_sam2[1:35]))

index_iesdash <- str_detect(cmoa_date$creation_date, patter = ies_dash)
cmoa_date[index_iesdash, 'creation_date'] <- iesdash_set

#there are years with the format of mid-1960s, I replace this instance with 1965
which(str_detect(cmoa_date$creation_date, pattern = "mid-1960s"))
cmoa_date[712, 'creation_date'] <- "1965"

#Remove other unncessary words such as centure, posthumous
word <- one_or_more(ANY_CHAR)
str_subset(cmoa_date$creation_date, pattern = word)

#Remove c.
c_dot <- "c" %R% DOT
sub_c_dot <- str_subset(cmoa_date$creation_date, pattern = c_dot) 
index_cdot <- which(str_detect(cmoa_date$creation_date, pattern = c_dot))
cmoa_date[index_cdot, 'creation_date'] <- str_replace(sub_c_dot, pattern = c_dot, replacement = "")

#Remove ca.
ca_dot <- "ca" %R% DOT
sub_ca_dot <- str_subset(cmoa_date$creation_date, pattern = ca_dot) 
index_cadot <- which(str_detect(cmoa_date$creation_date, pattern = ca_dot))
cmoa_date[index_cadot, 'creation_date'] <- str_replace(sub_ca_dot, pattern = ca_dot, replacement = "")

#remove /
slash<- dgt(4) %R% "/" %R% dgt(4)
slash_split <- str_split(slash_subset, pattern = "/")
slash_subset <- str_subset(cmoa_date$creation_date, pattern = slash)
numeric_slash <- sapply(slash_split, as.numeric, simplify = F)
mean_slash <- sapply(numeric_slash, mean, simplify = F)
round_slash <- sapply(mean_slash, round)
index_slash<- which(str_detect(cmoa_date$creation_date, pattern = slash))
cmoa_date[index_slash, 'creation_date'] <- round_slash

#remove space
space <- SPC
subset<- str_subset(cmoa_date$creation_date, pattern = space)
(str_extract(str_subset(cmoa_date$creation_date, pattern = space), pattern = dgt(4)))[!is.na(str_extract(str_subset(cmoa_date$creation_date, pattern = space), pattern = dgt(4)))]
which(str_detect(cmoa_date$creation_date, pattern = space))
index_im <- which(str_detect(str_subset(cmoa_date$creation_date, pattern = space), pattern = dgt(4)))
index_subset <- str_detect(cmoa_date$creation_date, pattern = space)

subset[index_im] <- (str_extract(str_subset(cmoa_date$creation_date, pattern = space), pattern = dgt(4)))[!is.na(str_extract(str_subset(cmoa_date$creation_date, pattern = space), pattern = dgt(4)))]
cmoa_date[index_subset, 'creation_date'] <- cmoa_date$updated_latest_year[index_subset]

#final sanity check
sum(str_detect(cmoa_date$creation_date, pattern = exactly(dgt(4))))
date <- select(cmoa_date, `title`, `creation_date`)
sum(is.na(cmoa_date$creation_date))
```
##The area of the photograph is more informative than its length or width alone. I use dplyr package to compute the area of the photograph
```{r}
cmoa_size <- select(cmoa, `title`, `item_width`,`item_height`,`item_depth`)
names(cmoa)
str(cmoa_size)
cmoa_size <- mutate(cmoa_size, surface = item_width * item_height )
```
##Full the join function from dplyr package to create masterlist, full artist table and etc
```{r}
#combine birthplace and deathplace of artist
bp_dp <- full_join(bp_combined, dp_combined)

#Combine birthplace, deathplace, birthdate and deathdate of the artist
full_artist <- full_join(bp_dp, bd.dd_combined)

#Create masterlist that combines both photographer and photograph information
cmoa_final <- cbind(date, cmoa_size[,c("item_width", "item_height", "item_depth", "surface")])
```
##Output spreadsheet and csv full
```{r}
write.xlsx(full_artist, "cmoa_cleaned1.xlsx", sheetName="full_artist", row.names = FALSE)
write.xlsx(bp_dp, "cmoa_cleaned1.xlsx", sheetName="birthp_as.matrideathp", row.names = FALSE, append = T)
write.xlsx(bp_combined, "cmoa_cleaned1.xlsx", sheetName="birthplace", row.names = FALSE, append = T)
write.xlsx(dp_combined, "cmoa_cleaned1.xlsx", sheetName="deathplace", row.names = FALSE, append = T)
write.xlsx(cmoa_size_trimmed, "cmoa_cleaned1.xlsx", sheetName="photosize", row.names = FALSE, append = T)
write.xlsx(cmoa_date, "cmoa_cleaned1.xlsx", sheetName="photodate", row.names = FALSE, append = T)
write.xlsx(cmoa_final, "cmoa_cleaned1.xlsx", sheetName="photoinfo", row.names = FALSE, append = T)
write.xlsx(masterlist, "cmoa_cleaned1.xlsx", sheetName="masterlist", row.names = FALSE, append = T)

write.csv(full_artist, "cmoa_fullartist.csv", row.names = F)
write.csv(bp_dp, "cmoa_bpdp.csv", row.names = F)
write.csv(masterlist, "cmoa_masterlist.csv", row.names = F)
```

